---
title: "Data Visualization and Analysis: Report"
author: 
- "Debanjali Biswas, Deepansh Pandey and Nikhil Acharya"
- "Computer Science Institute, Univeristy of Bonn"
bibliography: "/Users/debanjalibiswas/Uni Bonn/2nd sem/Data Analytics/Bike-Sharing-Dataset/Final_Project/bibliography.bibtex"
date: "July 29,2018"
output: pdf_document
fig_caption: yes
---

## Overview of Dataset

The dataset chosen is the Bike-Sharing-Dataset [@DataSet]. Bike sharing systems are new generation of traditional bike rentals where whole process from membership, rental and return 
back has become automatic. Through these systems, user is able to easily rent a bike from a particular position and return 
back at another position. Currently, there are about over 500 bike-sharing programs around the world which is composed of 
over 500 thousands bicycles. Today, there exists great interest in these systems due to their important role in traffic, 
environmental and health issues. 

Bike-sharing rental process is highly correlated to the environmental and seasonal settings. For instance, weather conditions, precipitation, day of week, season, hour of the day, etc. can affect the rental behaviors.

## Idea
To find the yearly sales of Bikes rented based on different factors which include weather conditions and type of users.

## Aproach
The daily records of the sales of bikes was loaded into a dataframe and was preprocessedin order to match the appropriate format for invidual attributes which was later used for plots and linear modeling. Pipeline functions from R was used mainly preprocessing and subsetting of data. Each of the subsets of the dataframe was used to create plots and the results were used to visualize the effect of these factors mentioned. The individual attrbutes were compared with the sales of bikes using Karl Pearson's co-effecient. Attributes with low correlation co-effecients were ignored. 

For predicting the sale of bikes for a given day, Linear Modeling was used. The results of the summary functions which include standard error, t-value, p-value and F-statistics, were used to choose the best attributes for the model. We also implemented a Linear Model and forecasted the result for Time Series Analysis.

## Understanding the Dataset
The following is a list of the attributes from the dataset:
```{r, echo=FALSE}
day_data <- read.csv('Bike-Sharing-Dataset/day.csv') # Reading the Dataset
names(day_data) # Prnting the attributes of the Dataset
```

## Bike Rental Trend
The trends (as shown in the Figure 1) below are depicts in that the count of Bike rentals increased in the month of March. It is also seen that this count increases with year.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2) # for data visulisation
library(plotly) # for plot rendering
library(dplyr) # for select statement
library(grid)
library(gridExtra)
library(date) # for handling dates
library(lubridate)
library(pacman)
library(forecast)
library(tidyr)
```

```{r fig1, fig.height=3, fig.align='center', fig.width=8, echo=FALSE, fig.cap='Monthly Bike Rental Trend'}
# Generates the Rental Trend
ggplot(data=day_data, aes(x=as.factor(mnth), y=cnt, fill= as.factor(year(dteday)))) + 
  geom_bar(stat="identity", position=position_dodge())+
  labs( x= "Month", y="Rental Count", fill= "Year") +
  scale_x_discrete(labels=month.abb) 

```


## Analysis of the Bike Rental count with respect to different factors of Weather
The different correlation co-effecients of the count of Bike Rentals with respect to each of the factors are as follows:
```{r, echo=FALSE}
print("With Humidity:")
cor(day_data$hum,day_data$cnt)
print("With Temperature:")
cor(day_data$cnt,day_data$temp)
print("With Real Feel Temperature:")
cor(day_data$cnt,day_data$atemp)
print("With Windspeed:")
cor(day_data$cnt,day_data$windspeed)
```

The real feel temperature largely influences the count of Bike rentals as it has the largest correlartion co-effecient.These effects are futher demonstrated in Figure 2.

```{r fig2, echo=FALSE, fig.height=4, fig.align='center', fig.width=8, fig.cap='Effect of different factors: Humidity,Temperature, Real feel Temperature, Windspeed'}
# Effect of Humidity on Bicycle Rentals
f1 <- ggplot(data=day_data, aes(x=hum, y=cnt)) +geom_point(size=0.00001)+ labs(x= "Humidity", y="Rental Count")  +scale_x_continuous(breaks=c(0,0.5,1))

#Effect of Temperature on Bike sales
f2 <- ggplot(data=day_data, aes(x=temp, y=cnt)) +geom_point(size=0.00001)+ labs(x= "Temperature", y="Rental Count")  +scale_x_continuous(breaks=c(0,0.5,1)) + geom_smooth(method='lm')

#Effect of Real feelTemperature on Bike sales
f3 <- ggplot(data=day_data, aes(x=atemp, y=cnt)) +geom_point(size=0.00001)+ labs(x= "Real feel Temperature", y="Rental Count")  +scale_x_continuous(breaks=c(0,0.5,1)) + geom_smooth(method='lm')

#Effect of windspeed on count of sales
f4 <- ggplot(data=day_data, aes(x=windspeed, y=cnt)) +geom_point(size=0.00001)+ labs(x= "Windspeed", y="Rental Count")  +scale_x_continuous(breaks=c(0,0.5,1)) + geom_smooth(method='lm')
grid.arrange(f1, f2, f3, f4, ncol=2, nrow=2)
```

## Analysis with respect to Other Weather Conditions
The count of Bike Rentals are the maximum when the sky is clear and the minimum when it is snowing or raining (as seen in Figure 3)

```{r fig3,echo=FALSE, fig.align='left', fig.width=6, fig.height=2, fig.cap='Effect of weather on rentals'}
stacked_plot_data = day_data %>% select(holiday,weekday,cnt,casual,registered,weathersit) %>%  
  mutate(public_holiday = ifelse(holiday == 0 ,"Non-Working Day", "Working Day"))
stacked_plot_data = stacked_plot_data %>% select(holiday,weekday,cnt,casual,registered,weathersit, public_holiday) %>% mutate(weather = case_when(weathersit==1 ~ "Clear" , weathersit==2 ~ "Mild" ,  weathersit==3 ~ "Snow/Rain"))

#Effect of weather on sales
ggplot(stacked_plot_data, aes(x = weather, y = cnt)) +
  geom_bar(stat = "identity") + scale_y_continuous(labels = scales::comma) + labs(x= "Weather Condition", y="Rental Count")
```
```{r fig4,echo=FALSE, fig.height=2, fig.align='right', fig.width=6, fig.cap='Rental Count per day of the week'}
#sales per day of the week based on holidays
ggplot() + geom_bar(aes(y = cnt, x = as.factor(weekday), fill = as.factor(public_holiday)), data = stacked_plot_data,stat="identity",position=position_dodge()) + labs( x= "Day", y="Rental Count" , fill="Holiday") + scale_y_continuous(labels = scales::comma)+ scale_x_discrete(labels= c("Sun", "Mon", "Tues", "Wed", "Thurs", "Fri", "Sat"))

```
## Analysis with respect to Casual and Registered users
The number of casual users renting is more during the weekends where the number of registered users renting is more during the weekdays (as in Figure 5). We also observe that count of rentals are more on Non-Working Days than Working Days. (Figure 4)

```{r fig5, echo=FALSE,fig.height=2, fig.align='center', fig.width=8, fig.cap='Effect of day on different users: Casual, Registered'}

#Count of casual users per day of the week
f1 <- ggplot(stacked_plot_data, aes(x = as.factor(weekday), y = casual, group=1))+ scale_x_discrete(labels= c("Sun", "Mon", "Tues", "Wed", "Thurs", "Fri", "Sat"))+ 
  geom_bar(stat = "identity") + scale_y_continuous(labels = scales::comma)+ labs(x= "Day", y="Casual Users Count")

#Count of registered users per day of the week
f2 <- ggplot(stacked_plot_data, aes(x = as.factor(weekday), y = registered, group=1)) + scale_x_discrete(labels= c("Sun", "Mon", "Tues", "Wed", "Thurs", "Fri", "Sat"))+
  geom_bar(stat = "identity") + scale_y_continuous(labels = scales::comma) + labs(x= "Day", y="Registered Users Count")

grid.arrange(f1, f2, ncol=2)
```

## Linaer Modeling
Linear modelling is done using the best parameters: real feel temperature, humidity, windspeed, season and weather. Summary of the model and the predicted value are shown below:

```{r,echo=FALSE}
#Linear modelling using best parameters real feel temperature, humidity, windspeed,season and weather. Explain why its a good fit
model <-lm(cnt~atemp+hum+windspeed+season+weathersit,data= day_data)
test_data = data.frame(atemp=0.165,hum=0.3,windspeed=0.4,season=1,weathersit=1)
summary(model)
#predict(model,test_data)

```

## Time Series Analysis
We are plotting the total count of bikes for various month. Implementing a linear model using lag function to do prediction and forecasting for out data. 

```{r,fig8,echo=FALSE, fig.height=2, fig.align='center', fig.width=8, fig.cap='Prediction by Linear Model'}

# We estimate the AR model using Linear Regression
# Here you can add something more about linear model
# Our dependent and independent variable
timeseries_model <- lm(cnt ~ lag(cnt) + 
              lag(cnt,2) +
              lag(cnt,3) +
              lag(cnt,4) +
              lag(cnt,5), data = day_data)

# We check for the standard error
# You can also describe other things just run the below line
#summary(timeseries_model)
#Here we are outputting our model predicted value and real values
day_data %>% transmute(dteday,cnt,prediction = c(0,0,0,0,0, predict(timeseries_model))) %>%
  gather("Type","Value",-dteday) %>%
  ggplot(aes(x = dteday, y = Value, color = Type)) + 
  labs(x='Date')+
  geom_point()
```
```{r,fig9,echo=FALSE, fig.height=3, fig.align='center', fig.width=8, fig.cap='Basic Forecasting'}

# we extract the seasonal effect and do basic forecasting
count <- day_data$cnt %>%
  ts(frequency = 365) %>%
  stl("periodic")

# here we use autoplot function 
# you can google more about these components of timeseries:
# 1.  Seasonality is the presence of variations that occur at specific regular intervals less than a year, such as weekly, monthly, or quarterly.
# 2.  Trend is the component of a time series that represents variations of low frequency in a time series, the high and medium frequency fluctuations having been filtered out.
# 3.  Remainder irregular component (or "noise") at time t, which describes random, irregular influences. It represents the residuals or remainder of the time series after the other components have been removed.
# Write the graph outcomes
#autoplot(count)
# Here we make a basic forcasting
plot(forecast(count)) 

```

## Conclusion
Humidity, Windspeed, Real feel Temperature, Seasons, other Weather conditions like snowfall, rain, etc. heavly influences the count of Bike rentals. The real feel temperature largely influences the count of Bike rentals as it has the largest correlartion co-effecient. Registered Users tend to use the bikes more on weekdays than weekends and non-working days while the casual users use it on the non-working days. 

## References


